---
title: "Netflix Content Performance Analysis"
author: "Jeremy Precel"
output:
  html_document:
    df_print: paged
---

# Overview

This project analyses a dataset of 100 Netflix titles to explore differences in audience reach, ratings and viewership consistency across content genres.

The analysis demonstrates data importing, cleaning, validation, transformation, joins and grouped descriptive analysis in R.

The dataset includes both categorical and numeric variables and contains the following fields:

- `show_id` — unique identifier for each title
- `title` — title name
- `type` — Movie or TV Show
- `release_year` — year of release
- `genre` — content genre
- `production_country` — country associated with the title
- `rating` — audience rating score
- `duration_min` — duration in minutes
- `views_millions` — estimated views in millions

The dataset used for this analysis contains **100 titles and 9 variables**.

# Setup

```{r setup, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(knitr)
```

# Import Data

The dataset was imported from a CSV file using `read_csv()`.

The original `country` variable was renamed to `production_country` to make the variable name more descriptive.

```{r import-data}
netflix_dataset <- read_csv("netflix_dataset.csv")

netflix_dataset <- netflix_dataset %>%
  rename(production_country = country)
```

# Data Inspection

The structure and dimensions of the dataset were inspected to confirm the number of observations, variables and data types.

```{r inspect-data}
str(netflix_dataset)

dim(netflix_dataset)

names(netflix_dataset)
```

The dataset contains **100 observations and 9 variables**.

Categorical variables were converted from character values to factors because they represent defined categories.

```{r convert-factors}
netflix_dataset$type <- as.factor(netflix_dataset$type)
netflix_dataset$genre <- as.factor(netflix_dataset$genre)
netflix_dataset$production_country <- as.factor(
  netflix_dataset$production_country
)

str(netflix_dataset)
```

# Data Quality Checks

The dataset was checked for missing values and category consistency.

```{r data-quality}
colSums(is.na(netflix_dataset))

levels(netflix_dataset$type)

levels(netflix_dataset$genre)

levels(netflix_dataset$production_country)
```

No missing values were identified in the dataset.

The dataset already follows a tidy structure:

- each row represents one title
- each column represents one variable
- each cell contains one value

No reshaping was required before analysis.

# Genre Lookup Table

A genre lookup table was created to assign a numeric identifier to each genre.

```{r genre-lookup}
genre_list <- list(
  genre = c(
    "Action",
    "Comedy",
    "Documentary",
    "Drama",
    "Romance",
    "Sci-Fi",
    "Thriller"
  ),
  genre_id = c(1, 2, 3, 4, 5, 6, 7)
)

genre_lookup <- data.frame(genre_list)

genre_lookup
```

The lookup table was joined to the main dataset using `left_join()`.

```{r join-data}
netflix_dataset <- left_join(
  netflix_dataset,
  genre_lookup,
  by = "genre"
)

head(netflix_dataset)
```

# Genre Performance Analysis

Grouped descriptive statistics were calculated to compare viewership and audience ratings across genres.

The analysis includes:

- number of titles
- mean viewership
- median viewership
- minimum viewership
- maximum viewership
- standard deviation of viewership
- average audience rating

```{r genre-summary}
genre_summary <- netflix_dataset %>%
  group_by(genre) %>%
  summarise(
    Count = n(),
    Mean_views = mean(views_millions),
    Median_views = median(views_millions),
    Minimum_views = min(views_millions),
    Maximum_views = max(views_millions),
    Standard_deviation = sd(views_millions),
    Average_rating = mean(rating)
  ) %>%
  arrange(desc(Mean_views))

genre_summary
```

# Average Viewership by Genre

The grouped analysis shows substantial differences in average audience reach across genres.

Documentary recorded the highest average viewership at approximately **155.05 million views**, closely followed by Sci-Fi at approximately **153.12 million**.

Drama recorded the lowest average viewership at approximately **85.99 million views**.

```{r viewership-table}
genre_summary %>%
  select(
    genre,
    Count,
    Mean_views,
    Median_views,
    Standard_deviation
  ) %>%
  kable(
    digits = 2,
    caption = "Viewership Performance by Genre"
  )
```

# Audience Ratings

Audience ratings were relatively similar across genres compared with the differences observed in viewership.

Thriller achieved the highest average rating at approximately **7.66**, despite ranking below several genres in average viewership.

Documentary generated the highest average viewership but had an average rating of approximately **6.95**.

```{r ratings-table}
genre_summary %>%
  select(
    genre,
    Mean_views,
    Average_rating
  ) %>%
  kable(
    digits = 2,
    caption = "Average Viewership and Audience Rating by Genre"
  )
```

# Viewership Variation

Standard deviation was used to compare how consistent viewership was within each genre.

Documentary recorded the lowest viewership variation at approximately **57.24 million**, indicating that viewership was comparatively consistent across the Documentary titles in this dataset.

Thriller recorded the highest variation at approximately **76.71 million**, indicating greater differences in audience size between Thriller titles.

```{r variation-table}
genre_summary %>%
  select(
    genre,
    Mean_views,
    Standard_deviation
  ) %>%
  arrange(Standard_deviation) %>%
  kable(
    digits = 2,
    caption = "Viewership Variation by Genre"
  )
```

# Key Findings

The analysis produced several notable findings.

**Audience Reach**

Documentary and Sci-Fi generated the highest average viewership within the dataset.

**Ratings vs Popularity**

The highest-rated genre was not the most-viewed genre. Thriller recorded the highest average audience rating, while Documentary recorded the highest average viewership.

This suggests that audience ratings and audience reach represent different dimensions of content performance.

**Content Consistency**

Documentary combined the highest average viewership with the lowest standard deviation, making it the most consistently viewed genre among the seven genres analysed.

# Scope

This analysis is based on a dataset of **100 Netflix titles**.

The findings should therefore be interpreted as observations from this sample and should not be treated as representative of Netflix's complete catalogue.